---
title: "genetic_correlations"
author: "<h3>Author</h3>Brian M. Schilder"
date: "`r Sys.Date()`"
output:
  BiocStyle::html_document:
vignette: >
    %\VignetteIndexEntry{phenomix} 
    %\usepackage[utf8]{inputenc}
    %\VignetteEngine{knitr::rmarkdown}
---

```{r setup, include=TRUE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, root.dir = here::here())

library(dplyr)
library(ggplot2)
library(googledrive)
library(data.table)
set.seed(2019)
```


Meta-analysis of genetic correlation studies between Parkinson's Disease and other phenotypes.

# Preprocessing

## DeGAs

```{r, eval=FALSE} 
### phenomix is currently under development (not yet released) ####
# devtools::install_github("neurogenomics/phenomix")
library(phenomix)

degas <- phenomix::get_DEGAS()
knn <- phenomix::find_neighors()
 
# data.table::fwrite(knn, here::here("data/DEGAS/DEGAS_contributionGene_corr.csv"))
```

## LD Hub 

I began to explore this dataset but then realized it does not contain any Parkinson's Disease GWAS.

```{r, eval=FALSE}
rg = xlsx::read.xlsx(here::here("data/LD_Hub/LD-Hub_genetic_correlation_221x221_no_ENIGMA.xlsx"), sheetName = "rG") %>% 
  tibble::column_to_rownames("NA.")
colnames(rg) <- gsub("[.]","-",colnames(rg))
row.names(rg) <- gsub("[.]","-",row.names(rg))
rg[rg=="/"] <- NA 
rg <- Matrix::as.matrix(rg, sparse = T)


rp = xlsx::read.xlsx(here::here("data/LD-Hub_genetic_correlation_221x221_no_ENIGMA.xlsx"), sheetName = "rP")%>%
    tibble::column_to_rownames("NA.")
colnames(rp) <- gsub("[.]","-",colnames(rp))
row.names(rp) <- gsub("[.]","-",row.names(rp))
rp[rp=="/"] <- NA 
rp <- Matrix::as.matrix(rp, sparse = T)

```

Check for Parkinson's

```{r, eval=FALSE}
grep("Parkinson",colnames(rp), ignore.case = T, value = T)
```



## Meta-analysis data

```{r}
googledrive::drive_download("https://docs.google.com/spreadsheets/d/19jz9l2P7W2f1PWT9t3x0L8h6VQL021RdRI7D4FAOjX0/edit?usp=sharing#gid=1366974120", path = here::here("data/metaanalysis/TableS2.xlsx"), overwrite = TRUE)

sheets <- readxl::excel_sheets(here::here("data/metaanalysis/TableS2.xlsx")) 
```

## Merge datasets

```{r}
corr_sheets <- sheets[endsWith(sheets,"_corr")]
 
corr_data <- lapply(corr_sheets, function(sheet){
  print(sheet)
  dat <- readxl::read_excel(here::here("data/metaanalysis/TableS2.xlsx"), sheet = sheet) 
  dat$Sheet <- gsub("_corr","",sheet) 
  cols <- c("Trait1","Trait2","corr",
            "Reference","Source","Dataset")  
  cols <- cols[cols %in% colnames(dat)]
  return(dplyr::relocate(dat, all_of(cols), .before = 1))
}) %>% `names<-`(corr_sheets) %>%
  data.table::rbindlist(fill = TRUE)

```

## Harmonise names

Attempt to somewhat harmonise phenotype names across studies.

```{r}
fix_str <- function(vec){
  vec %>% 
    tolower() %>%
    gsub(pattern = "[ ]|-|[.]|[:]|[\\]|[\\/]",replacement = "_") %>%
    gsub(pattern = "\'|\"|[(]|[)]",replacement = "") %>%
    gsub(pattern = "_+",replacement = "_") %>%
    # stringr::str_wrap(width = 50) %>%
    stringr::str_trunc(width = 50) 
}

corr_data <- corr_data %>% 
  dplyr::mutate(Trait1 = fix_str(Trait1),
                Trait2 = fix_str(Trait2))
```



# Top traits

Get the top N traits per study that are most highly correlated with Parkinson's Disease.

## Gather 

```{r}
max_traits <- 10
  
top_corr <- corr_data %>% 
  dplyr::group_by(Sheet) %>%
  subset(Trait1!=Trait2) %>%
  #### Correct for the number of correlations ? ####
  dplyr::mutate(corr_q = abs(corr)/n(),
                corr_z = scale(abs(corr))[, 1],
                valence = ifelse(corr>0,"+",ifelse(corr<0,"-",0)),
                n_tests = n()) %>%
  dplyr::ungroup() %>%
  dplyr::group_by(Sheet, Trait1) %>%
  dplyr::mutate(id1 = paste0(Trait1,"@",dplyr::group_indices()),
                id2 = paste0(Trait2,"@",dplyr::group_indices())) %>%
  dplyr::slice_max(order_by = corr_q,
                   n = max_traits, 
                   with_ties = TRUE) %>%
  dplyr::select(Sheet, Trait1, Trait2, id1, id2, 
                corr, corr_q, corr_z, valence, Reference, Source, n_tests) %>% 
  dplyr::ungroup() %>%
  # dplyr::arrange(dplyr::desc(abs(corr_q))) %>%
  data.table::data.table()
top_corr$Trait2 <- factor(x = top_corr$Trait2,
                          levels = rev(unique(top_corr$Trait2)),
                          ordered = TRUE)
knitr::kable(top_corr)

```

### Save merged data

```{r}
data.table::fwrite(x = corr_data[,c("Sheet","Trait1","Trait2","corr","Reference","Source")],
                   file = here::here("data/metaanalysis","merged_correlations.csv"))
```



## Plot

### Bar plot

```{r}
gg_cor <- ggplot(data = top_corr, aes(x=corr, y = Trait2, fill = corr)) +
  geom_col() +
  geom_point(show.legend = FALSE, alpha=.5) +
  facet_grid(facets = paste0(gsub("[(]","\n(",Sheet),
                             "\n (tests = ",formatC(n_tests,big.mark = ","),")") ~., 
             space = "free",
             scales = "free",
             drop = FALSE) +
  scale_x_continuous(limits = c(-1,1)) +
  scale_y_discrete(labels = setNames(gsub("@.*","",levels(top_corr$Trait2)),
                                     levels(top_corr$Trait2) ) ) +
  scale_fill_gradient(low = "red", high = "blue", limits = c(-1,1)) +
  labs(title = "Genetic correlations with Parkinson's Disease",
       x = "Correlation") + 
  theme_bw() +
   theme(strip.background = element_rect(fill = "black"), 
        strip.text = element_text(color = "white", angle = 0),
        strip.text.y = element_text(color = "white", angle = 0), 
        panel.grid.minor.x = element_blank())
print(gg_cor)

ggplot2::ggsave(filename = here::here("plots/metaanalysis/correlation_metaanalysis.pdf"),
                plot = gg_cor, dpi = 400, height = 8, width = 9)
ggplot2::ggsave(filename = here::here("plots/metaanalysis/correlation_metaanalysis.png"),
                plot = gg_cor, dpi = 400, height = 8, width = 9)
```

# Session Info

<details> 

```{r}
utils::sessionInfo()
```

</details>

