---
title: "geneshot"
author: "<h3>Author</h3>Brian M. Schilder"
date: "`r Sys.Date()`"
output:
  BiocStyle::html_document:
vignette: >
    %\VignetteIndexEntry{phenomix} 
    %\usepackage[utf8]{inputenc}
    %\VignetteEngine{knitr::rmarkdown}
---

```{r setup, include=T, warning=F, message=F}
knitr::opts_chunk$set(echo = TRUE, root.dir=here::here("/Desktop/PD_omics"))
# knitr::opts_knit$set(echo = TRUE, root.dir=here::here("/Desktop/PD_omics"))

library(dplyr)
library(tidyr)
library(ggplot2)
set.seed(2019)
```

# Geneshot

Plot genes that come from entering "Parkinson's" into [Geneshot](https://amp.pharm.mssm.edu/geneshot/index.html?searchin=Parkinson's&searchnot=&rif=autorif), which gathers associated gene names from PubMed literature.  

```{r}
replace_ENSP <- function(PD_genes,
                         gene_col="Gene"){ 
  inds <- which(startsWith(PD_genes[[gene_col]], "ENSP"))
  if(length(inds)>0){
    ensembl_ids <- PD_genes[[gene_col]][inds]
    conversion <- AnnotationDbi::mapIds(EnsDb.Hsapiens.v75::EnsDb.Hsapiens.v75,
                                        keys = ensembl_ids,
                                        keytype = "PROTEINID",
                                        column = "SYMBOL")
    PD_genes[inds,gene_col] <- unname(conversion)
    PD_genes <- PD_genes[!is.na(PD_genes[[gene_col]]),]
  } 
  return(PD_genes)
}
```

## Associated genes  

```{r Associated genes }
PD_genes <- data.table::fread(here::here("data/Geneshot/geneshot.parkinsons_disease.tsv")) %>% 
  dplyr::mutate(Zscore=scales::rescale(`Fraction of publications from total gene publication`)) %>%
  dplyr::select(Rank, Gene, Zscore) 
PD_genes <- replace_ENSP(PD_genes)
```


## Predicted genes 

```{r}
gene_counts <- data.table::fread(here::here("data/Geneshot/geneshot.parkinsons_disease.tsv"))
PD_pred_genes <- data.table::fread(here::here("data/Geneshot/geneshot_predictions_AutoRIF.parkinsons_disease.tsv")) %>%
  dplyr::mutate(Zscore=scales::rescale(Score),
                Type="Geneshot") %>%  
  dplyr::select(ID=Rank,Name=Gene,Zscore,
                Genes="Genes contributing most to score:similarity score",
                Type) %>%
  dplyr::mutate(Type_ID=paste(Type,ID,sep="_")) %>% 
  merge(gene_counts, all.x = T, by.x = "Name", by.y = "Gene")

PD_pred_genes <- replace_ENSP(PD_pred_genes, gene_col = "Name")
  
PD_pred_genes_melt <- PD_pred_genes %>%
  tidyr::separate(col="Genes", 
                  sep = ";",
                  into = paste0("gene",1:9)) %>% 
  data.table::melt.data.table(id.vars = c("ID","Name","Zscore","Type","Type_ID","Rank","Publication count",
                                          "Fraction of publications from total gene publication")) %>% 
  tidyr::separate(col="value",sep = ":", into=c("gene2","similarity")) %>%
  subset(!is.na(gene2))


saveRDS(PD_pred_genes_melt,here::here("data/Geneshot/PD_pred_genes_melt.rds"))
```

# Geneshot UMAP

[Density plots tutorial](https://www.r-graph-gallery.com/2d-density-plot-with-ggplot2.html)

## Constuct matrix 

```{r}

EXP <- reshape2::dcast(PD_pred_genes_melt,
                       formula = Name ~ gene2, 
                       value.var = "similarity", 
                       fill = 0) %>%
  tibble::column_to_rownames("Name")  %>%
  as.matrix() %>%
  as("sparseMatrix")
```



## Run UMAP

```{r Geneshot UMAP}
library(network)
library(ggnetwork)
library(knn.covertree)
 

# seurat <- Seurat::CreateSeuratObject(as(t(EXP), "sparseMatrix"), 
#                                      meta.data = data.frame(PD_pred_genes,
#                                                             row.names = PD_pred_genes$Name))
# seurat <- Seurat::NormalizeData(seurat)
# seurat <- Seurat::FindVariableFeatures(seurat)
# seurat <- Seurat::ScaleData(seurat)
# seurat <- Seurat::RunPCA(seurat, npcs=50)
# seurat <- Seurat::FindNeighbors(seurat, force.recalc = T)
# seurat <- Seurat::RunUMAP(seurat, dims=1:20)
# seurat <- Seurat::FindClusters(seurat)
# Seurat::DimPlot(seurat, group.by = "Zscore") + Seurat::NoLegend()

#### UMAP ####
umap_res <- uwot::umap(X = as.matrix(EXP), n_components = 2,
                       ret_extra = c("model","nn","fgraph"))# %>% data.frame() 
umap_df <- umap_res$embedding %>% 
  `colnames<-`(paste0("UMAP",1:ncol(umap_res$embedding)))

#### KNN ####
knn_res <- knn.covertree::find_knn(umap_df, k=3)
# dim(knn_res$dist_mat)
adjGraph <- igraph::graph_from_adjacency_matrix(knn_res$dist_mat,#umap_res$fgraph, 
                                                mode = "undirected", 
                                                weighted = T)
clust <- igraph::cluster_louvain(adjGraph)


```
## Plot UMAP

### 2D

```{r} 
#### Annotated UMAP ####
umap_dat <- merge(PD_pred_genes, 
                  cbind(data.frame(umap_df),
                        Name = row.names(EXP),
                        cluster=clust$membership),
                  by="Name") %>% 
  dplyr::arrange(desc(Zscore)) 

 
label_genes <- umap_dat %>% 
  dplyr::group_by(cluster) %>%
  dplyr::slice_head(n = 5)

#### plot ####
gg_umap <- ggplot(umap_dat, aes(x=UMAP1, y=UMAP2, color=Zscore, 
                                # size=`Publication count`, 
                                size=Zscore*2,
                                label=Name)) +   
  stat_density_2d(aes(fill = ..level..), geom = "polygon",adjust = .2, contour_var = "ndensity") +
   scale_fill_distiller(palette=4, direction=-1) +
  scale_x_continuous(expand = c(0.1, 0.1)) +
  scale_y_continuous(expand = c(0.1, 0.1)) + 
   # stat_density_2d_filled(show.cegend = F, contour_var = "ndensity",
   #                       # linejoin = "bevel",
   #                       # alpha=.8,
   #                       breaks = seq(0.1, 1.0, length.out = 19)) +
    # scale_fill_manual(values = c("black",pals::inferno(n = 18))) +
   geom_point(alpha=.8) + 
  scale_color_gradientn(colours = c("magenta","blue")) +
  ggrepel::geom_label_repel(data = label_genes, max.overlaps = 20,na.rm = T,
                            aes(x=UMAP1, y=UMAP2, label=Name),
                            fill = alpha(c("white"),0.9),
                            seed = 2019,
                            show.legend = F) +
  theme_void() + 
   theme(plot.background = element_rect(fill = "black"),
         panel.background = element_rect(fill = "black"),
         legend.text = element_text(color="white"),
         legend.title = element_text(color="white"))  +
  guides(size=guide_legend(override.aes=list(fill="white")))

print(gg_umap)
 
#### Save ####
# ggsave(here::here("plots/geneshot_UMAP.jpg"), gg_umap, dpi = 300, height = 8, width = 10)
# ggsave(here::here("plots/geneshot_UMAP.pdf"),gg_umap, dpi = 300, height = 8, width = 10)

```
### 2D interactive

```{r}
plotly::ggplotly(gg_umap)
```


### 3D

Except this is like the one case where [Rayshader doesn't work...](https://github.com/tylermorganwall/rayshader/issues/148)

```{r 3D, eval=FALSE}
library(rayshader) 

rgl::clear3d()
plot_gg(gg_umap, width = 5, height = 5,
        multicore = TRUE, scale = 250, 
        zoom = 0.7, theta = 10, phi = 30, windowsize = c(800, 800))
```

## Similarity networks 

Experimented with similarity networks of gene-gene data. A bit unncessary since UMAP constructs networks anyways.

```{r, eval=FALSE}
library(BUS)

#### Construct similarity matrix using correlation (corr) or mutual information (MI)
sM <- BUS::gene.similarity(EXP, measure="corr",net.trim="none")
pM <- BUS::gene.pvalue(EXP, measure="corr",net.trim="none") 
netM <- BUS::pred.network(pM, sM)
netM[netM==0] <- NA

#### Alternative plotting method with iGraph ####
# adjGraph <- igraph::graph_from_adjacency_matrix(sM, weighted = T, mode = "undirected")
# clusts <- igraph::cluster_fast_greedy(adjGraph)
# plot(adjGraph)

#### Construct network object ####
edges <- reshape2::melt(netM) %>% `colnames<-`(c("from","to","weight"))
nodes <- predictions[,c("Gene","Rank","Score")]
net <- network(edges, vertex.attr = nodes, matrix.type = "edgelist", ignore.eval = F) 
# help(layout)

netF <- ggnetwork(net,  
                  layout = "fruchtermanreingold",
                  cell.jitter = 0)
netF <- netF[!is.na(netF$weight),]
netF <- netF[netF$weight>.005,]

ggplot(netF, 
       aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_edges(aes(alpha=weight), color = "black", na.rm = T) +
  geom_nodes(aes(size=Score), color = "grey") +
  geom_nodetext(aes(color = Score, label = Gene, size=Score),
                fontface = "bold", nudge_y = .05) +
  theme_blank()
```

# Wordcloud

Turn genes into word cloud.

[`ggwordcloud` tutorial](https://cran.r-project.org/web/packages/ggwordcloud/vignettes/ggwordcloud.html)

```{r Wordcloud}
library(ggwordcloud) 

ggplot(PD_genes[1:100,], aes(label = Gene, size = Zscore, color=Zscore)) +
  geom_text_wordcloud_area() + 
  theme_minimal()
```

  

# Session info 
 
<details>

```{r}
utils::sessionInfo()
```

</details> 

