---
title: "metaanalysis_merge" 
author: "<h3>Author</h3>Brian M. Schilder"
date: "`r Sys.Date()`"
output:
  BiocStyle::html_document:
vignette: >
    %\VignetteIndexEntry{phenomix} 
    %\usepackage[utf8]{inputenc}
    %\VignetteEngine{knitr::rmarkdown}
---

```{r setup, include=TRUE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, root.dir = here::here())
# knitr::opts_knit$set(echo = T, root.dir = here::here())

library(dplyr)
library(ggplot2)
library(googledrive)
library(patchwork)
set.seed(2019)

#### Import utility functions ####
source(here::here("code/utils.R"))
```


Merge all cell-type and tissue enrichment datasets for meta-analysis.


# Merge cell-types   


## Download 

You can download the meta-analysis tables from Google Drive and process them however you like.

```{r, eval=FALSE} 
# Table 1
googledrive::drive_download("https://docs.google.com/spreadsheets/d/18bKV3mUuyIv1yPN7gps9HPmhROb4jz0x7KxtA8mf0iM/edit#gid=743954199", path = here::here("data/metaanalysis/TableS1.xlsx"), overwrite = TRUE)

sheets <- readxl::excel_sheets(here::here("data/metaanalysis/TableS1.xlsx")) 
```

## Automated merging 
 
Alternatively, you can use the function `merge_data` we've provided in 
*code/utils.R* to automatically download, merge and harmonise the data.

```{r}
celltype_data <- merge_data(googledrive_url = "https://docs.google.com/spreadsheets/d/18bKV3mUuyIv1yPN7gps9HPmhROb4jz0x7KxtA8mf0iM/edit#gid=743954199",
                            file_name = "TableS1.xlsx",
                            sheet_search = "celltypes")
```


## Preprocess  

### FDR plots
 
```{r}
plot_celltypes <- celltype_data %>%    
  subset(!is.na(Sig) & !is.na(Cell_Type) & !( Cell_Type %in% c("brain"))) %>%
  dplyr::mutate( Sig_neglog =  -log1p(Sig)) %>% 
  dplyr::group_by(Reference, .drop = F) %>%
  dplyr::mutate(Cell_Type= paste0(Cell_Type,"@",group_indices())) %>%
  ##  Calculate zscore for fine-mapping overlap   
  dplyr::summarise(Cell_Type,Source,Sheet,Dataset,Sig,Count, 
                   FDR = p.adjust(p = Sig, method = "fdr"),
                   bonf = p.adjust(p = Sig, method = "bonf"),
                   n_tests = n()) %>% 
  dplyr::mutate(Score = ifelse(Reference=="Schilder et al, 2020 (bioRxiv)", 
                               as.numeric(cut(Count, 10))/10, 
                               scales::rescale(-log10(FDR), c(0.0000000001,1))) ) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(FDR_all = p.adjust(p = Sig, method = "fdr")) %>%
  subset(FDR_all<.05 & FDR<.05 & Sig < 0.05  & (!is.infinite(Score)))   
# Sort by medians 
plot_celltypes <- plot_celltypes %>% 
  dplyr::group_by(Reference, Cell_Type) %>%
  dplyr::summarise(mean_Score = mean(Score, na.rm=T),
                   median_Score = median(Score, na.rm=T),
                   .groups = "keep") %>%
  merge(plot_celltypes, by=c("Reference","Cell_Type"))  %>%
  dplyr::arrange(Reference, desc(median_Score), bonf) 
plot_celltypes$Cell_Type <- factor(plot_celltypes$Cell_Type, 
                                   levels = rev(unique(plot_celltypes$Cell_Type)), ordered = T)
                 
gg_FDR1 <- gg_FDR2 <- plot_fdr(dat = plot_celltypes, y="FDR", label = "Cell_Type")
plotly::ggplotly(gg_FDR1)

gg_FDR2 <- plot_fdr(dat = plot_celltypes, label = "Cell_Type")
plotly::ggplotly(gg_FDR2)
```

### Cell-types plot data

```{r}
createDT(plot_celltypes)
```


### Cell-types counts

```{r}
celltype_counts <- plot_celltypes %>% dplyr::group_by(Reference) %>% 
  dplyr::summarise(n_celltypes=dplyr::n_distinct(Cell_Type)) %>%
  dplyr::arrange(desc(n_celltypes))
knitr::kable(celltype_counts)
```


## Summary plot

Summarise the top celltypes/tissues per study according to abstracted "Score" value (computed here).

```{r, fig.height=10}  
### Limit the number of celltypes/study for visualization purposes
max_celltypes <- 25

celltype_plot <- ggplot(plot_celltypes %>% 
                          dplyr::group_by(Reference) %>% 
                          dplyr::slice_max(order_by = median_Score,
                                           n = max_celltypes),
       aes(x=Score, y=Cell_Type, fill=Reference)) + 
  # geom_bar(show.legend = F, stat="identity") +
  geom_boxplot(show.legend = FALSE) +
  # geom_violin(show.legend = F) +
  geom_point(show.legend = FALSE, alpha=.2) + 
  # Create a dicionary mapping the new labels
  scale_y_discrete(labels = setNames(gsub("@.*","",levels(plot_celltypes$Cell_Type )),
                                     levels(plot_celltypes$Cell_Type) ) ) + 
  # facet_wrap(facets = .~paste(gsub("[(]","\n(",Reference),"\n\n n tests =",n_tests),
  #          scales = "free", drop = F, nrow = 1) +  
  facet_grid(facets = paste0(gsub("[(]",
                                  "\n(",Sheet),
                             "\n (tests = ",
                             formatC(n_tests,big.mark = ","),")")~.,
           scales = "free", space="free", drop = F) + 
  theme_bw() +
  theme(strip.background = element_rect(fill = "black"), 
        strip.text = element_text(color = "white", angle = 0),
        strip.text.y = element_text(color = "white", angle = 0),
         panel.grid.minor.x = element_blank())

print(celltype_plot)
 
# ggsave(here::here("plots/metaanalysis/celltype_metaanalysis.pdf"),celltype_plot, dpi = 300, height = 15)
```





# Merge tissues 

## Automated merging 

```{r}
tissue_data <- merge_data(googledrive_url = "https://docs.google.com/spreadsheets/d/18bKV3mUuyIv1yPN7gps9HPmhROb4jz0x7KxtA8mf0iM/edit#gid=743954199",
                            file_name = "TableS1.xlsx",
                            sheet_search = "tissues")
```

## Preprocess  

### FDR plots

```{r}
plot_tissues <- tissue_data %>%   
  subset(!(Tissue %in% c(NA,"NA")) & (!is.na(Sig))) %>%
  dplyr::mutate(Sig_neglog =  -log1p(Sig)) %>% 
  ##  Calculate zscore for fine-mapping overlap  
  dplyr::group_by(Reference, .drop = FALSE) %>%
  dplyr::mutate(Tissue= paste0(Tissue,"@",group_indices())) %>%
  dplyr::summarise(Tissue,Source,Sheet,Dataset,Sig,PP.H4,  
                   FDR = p.adjust(p = Sig, method = "fdr"),  
                   bonf = p.adjust(p = Sig, method = "bonf"), 
                   n_tests = n(),
                   .groups = "keep") %>% 
  # dplyr::ungroup() %>%
  dplyr::mutate(Score = ifelse(Reference=="Schilder et al, 2020 (bioRxiv)", 
                               as.numeric(cut(PP.H4, 10))/10, 
                               scales::rescale(-log(FDR), c(0.0000000001,1))) ) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(FDR_all = p.adjust(p = Sig, method = "fdr")) %>%
  subset((FDR_all<.05 & FDR<.05 & Sig < 0.05) | (PP.H4>.8)) 
# Sort by medians 
plot_tissues <- plot_tissues %>%  
  dplyr::group_by(Reference, Tissue, .drop = F) %>% 
  dplyr::summarise(mean_Score = mean(Score, na.rm=T),
                   median_Score = median(Score, na.rm=T),
                   n_tissues = n_distinct(Tissue),
                   .groups = "keep") %>%
  merge(plot_tissues, by=c("Reference","Tissue"), all.y = T)  %>%
  dplyr::arrange(Reference, desc(median_Score), bonf) 
plot_tissues$Tissue <- factor(plot_tissues$Tissue,
                              levels = rev(unique(plot_tissues$Tissue)),
                              ordered = TRUE)
  
 
gg_FDR1 <- plot_fdr(dat = plot_tissues, y="FDR", label = "Tissue")
plotly::ggplotly(gg_FDR1)

gg_FDR2 <- plot_fdr(dat = plot_tissues, label = "Tissue")
plotly::ggplotly(gg_FDR2) 
```

### Tissues plot data

```{r}
createDT(plot_tissues)
```

### Tissues counts

```{r}
tissue_counts <- plot_tissues %>% dplyr::group_by(Reference) %>% 
  dplyr::summarise(n_tissues=dplyr::n_distinct(Tissue)) %>%
  dplyr::arrange(desc(n_tissues))
createDT(tissue_counts)
```


## Summary plot

```{r, fig.height=10, dpi=300, warning=FALSE}
# plot_tissues[is.na(plot_tissues$Dataset),"Dataset"] <- plot_tissues[is.na(plot_tissues$Dataset),"Sheet"]
max_tissues <- 15
                          

tissue_plot <- ggplot(plot_tissues %>%
                        dplyr::group_by(Reference) %>% 
                          dplyr::slice_max(order_by = median_Score,
                                           n = max_tissues), 
       aes(x=Score, y=Tissue, fill=Reference)) + 
  # geom_bar(show.legend = F, stat="identity") +
  geom_boxplot(show.legend = FALSE) +
  # geom_violin(show.legend = F) +
  geom_point(show.legend = FALSE, alpha=.2) +
  # facet_wrap(facets = .~paste(gsub("[(]","\n(",Reference),"\n\n n tests =",n_tests),
  #          scales = "free", drop = F, nrow = 1) +  
  facet_grid(facets = paste0(gsub("[(]","\n(",Sheet),
                             "\n (tests = ",
                             formatC(n_tests,big.mark = ","),")")~.,
           scales = "free", space="free", drop = FALSE) + 
  scale_x_continuous(limits = c(0,1)) +
  # Create a dicionary mapping the new labels
  scale_y_discrete(labels = setNames(gsub("@.*","",levels(plot_tissues$Tissue)),
                                     levels(plot_tissues$Tissue) ) ) + 
  theme_bw() +
  theme(strip.background = element_rect(fill = "black"), 
        strip.text = element_text(color = "white", angle = 0),
        strip.text.y = element_text(color = "white", angle = 0), 
        panel.grid.minor.x = element_blank())

print(tissue_plot)

# ggsave("plots/metaanalysis/tissue_metaanalysis.pdf",tissue_plot, dpi = 300, height = 15)
```

# Fuse plots

```{r, fig.height=12, fig.width=15, dpi=300}
fused_plot <- (celltype_plot + labs(title="Cell-types") + 
                 tissue_plot + labs(title="Tissues")) +
  patchwork::plot_annotation(tag_levels = letters)
print(fused_plot)

ggsave(here::here("plots/metaanalysis/metaanalysis.pdf"),fused_plot, height = 14, width = 15, dpi = 400)
ggsave(here::here("plots/metaanalysis/metaanalysis.jpg"),fused_plot, height = 14, width = 15, dpi = 400)
```


# Session Info

<details> 

```{r}
utils::sessionInfo()
```

</details>
