---
title: "multimodal_networks"
author: "<h3>Author</h3>Brian M. Schilder"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_document: 
    theme: spacelab
    highlight: zenburn 
    code_folding: show 
    toc: true 
    toc_float: true
    smooth_scroll: true
    number_sections: false 
    self_contained: true 
---

```{r setup, include=T, warning=F, message=F}
knitr::opts_chunk$set(echo = TRUE, root.dir=here::here("/Desktop/PD_omics"))
# knitr::opts_knit$set(echo = TRUE, root.dir=here::here("/Desktop/PD_omics"))

library(dplyr)
library(ggplot2)
library(orthogene)
set.seed(2019)
```

 
# Cell-types / Tissues/ Pathways 

Use the genes from GeneShot to then check for similarity with other term-associated gene sets.

## [CellMeSH](https://uncurl.cs.washington.edu/db_query) 

### Preprocess data

```{r CellMeSH}
cellmarker <- data.table::fread(here::here("data/UNCURL/PD_terms/cellmarker.csv"), select = 1:4) %>% 
  dplyr::mutate(Zscore=scales::rescale(-log(`P-value`),c(0,1))) %>%
  dplyr::select(ID=`Cell Ontology ID`, Name=`Cell`, Zscore, Genes=`Overlapping Genes`)%>%
  dplyr::mutate(Type="CellMarker")


cellmesh <- data.table::fread(here::here("data/UNCURL/PD_terms/cellmesh.csv"), select = 1:4) %>% 
  dplyr::mutate(Zscore=scales::rescale(`Log-likelihood`,c(0,1))) %>%
  dplyr::select(ID=`MeSH ID`, Name=`Cell Name`, Zscore, Genes=`Overlapping Genes`)%>%
  dplyr::mutate(Type="CellMeSH")

cellmesh_anatomy <- data.table::fread(here::here("data/UNCURL/PD_terms/cellmesh-anatomy.csv"), select = 1:4) %>%
  dplyr::mutate(Zscore=scales::rescale(`Log-likelihood`,c(0,1))) %>%
  dplyr::select(ID=`MeSH ID`, Name=`Cell Name`, Zscore, Genes=`Overlapping Genes`) %>%
  # Lots of celltypes in the "anatomy" section
  subset(!grepl("neuron|astrocyte|fibroblast|epithelial|endothelial|macrophages|monocytes|lymphocytes|leukocytes",
                tolower(Name))) %>%
  dplyr::mutate(Type="Anatomy")

GO <- data.table::fread(here::here("data/UNCURL/PD_terms/geneontology.csv"), select = 1:4) %>%
  dplyr::mutate(Zscore=scales::rescale(-log10(FDR),c(0,1))) %>%
  subset(FDR<.05) %>%
  dplyr::select(ID=`GO ID`, Name, Zscore, Genes=`Overlapping Genes`) %>%
  subset(!grepl("neuron|dendrite|axon",tolower(Name))) %>%
  dplyr::mutate(Type="GO")

annot <- rbind(cellmarker, cellmesh, cellmesh_anatomy, GO) %>% 
   dplyr::mutate(Type_ID=paste(Type, ID, sep="_"))%>% 
   dplyr::arrange(desc(Zscore)) %>% 
   dplyr::group_by(Type) %>%  
   dplyr::slice_head(n = 20) %>%
  data.table::data.table()
annot %>% dplyr::group_by(Type) %>% count()

{
  annot_genes <- tidyr::separate(annot,col = "Genes", sep = ", ", 
                               into=paste0("gene",1:100)) %>%
  data.table::melt.data.table(id.vars = c("Type_ID","Type","ID","Name","Zscore"), 
                              value.name = "Gene")  %>%
    subset(Type=="GO" & !is.na(Gene)) %>%
    dplyr::select(-variable)
    
  gene_map <- orthogene::map_genes(genes = annot_genes$Gene, 
                                   mthreshold = 1,
                                   drop_na = TRUE) 
  gene_dict <- setNames(gene_map$name, gene_map$input)
  annot_genes$Gene <- gene_dict[annot_genes$Gene] 
} 
```
### Import Geneshot data

Import data produced in `geneshot.Rmd`.

```{r}
PD_pred_genes_melt <- readRDS(here::here("data/Geneshot/PD_pred_genes_melt.rds"))
```



### Integrate all levels 

```{r Integrate all levels }
#### Merge with gene-level data ####
annot_fused <- rbind(annot_genes, 
                     PD_pred_genes_melt[1:100,] %>% 
                       dplyr::mutate(Type="Geneshot",) %>%
                       dplyr::select(Type, ID, Name, Zscore, Gene=gene2, Type_ID)
                     ) %>% 
  dplyr::mutate(Type_ID=paste(Type,ID,sep="_"),
                pseudo=1) %>% 
  # subset(Type!="Geneshot") %>%
  subset(!is.na(Gene))
data.table::fwrite(annot_fused, here::here("data/UNCURL/PD_terms/annot_fused.csv"))

### 
annot_unique <- annot_fused %>% 
  dplyr::select(-Gene) %>% unique() 
  
annot_mat <- data.table::dcast(data = annot_fused, 
                               formula = Type_ID  ~ Gene, 
                               value.var = "Zscore", 
                               fun.aggregate = mean,
                               fill = 0) %>% 
  tibble::column_to_rownames("Type_ID") 

data.table::fwrite(annot_mat, here::here("data/UNCURL/PD_terms/annot_fused.mat.csv"), row.names = TRUE)
```


## Multi-modal network

Now let's connect these terms at multiple levels using the gene names.

[Various network plotting methods](https://www.jessesadler.com/post/network-analysis-with-r).


# Multimodal wordcloud

```{r Multimodal wordcloud}
library(ggwordcloud)

annot <- rbind(cellmarker, cellmesh, cellmesh_anatomy, GO[1:20,] )
annot$Type <- factor(annot$Type, levels = c("CellMarker","CellMeSH","Anatomy","GO","Geneshot"),
                     ordered = T)

gg_annot <- ggplot(annot, aes(label = Name, size = Zscore, color=Type,  x=Type)) +
  geom_text_wordcloud_area() + 
  theme_minimal()
print(gg_annot)
```

# Plot PD annotations

Plot the PD annotations (Tissues, celltypes, GO). 

## Run UMAP

```{r}
umap_res <- uwot::umap(X = annot_mat, 
                       n_components = 5, 
                       ret_extra = c("model","nn","fgraph"))
##### high dimensional fuzzy graph ####
umap_dist <- as.matrix(umap_res$fgraph) %>% 
  `row.names<-`(row.names(annot_mat)) %>%
  `colnames<-`(row.names(annot_mat))
```

## Plot UMAP

```{r} 
#### Embeddings  ####
umap_df <- data.frame(umap_res$embedding,  
                      row.names = row.names(annot_mat) ) %>%
  `colnames<-`(paste0("UMAP",1:ncol(umap_res$embedding))) %>%
  tibble::rownames_to_column("Type_ID") %>%
  merge(annot_unique, by="Type_ID")


## Labels 
umap_labels <- umap_df %>% 
  # dplyr::select(-Gene, -variable) %>% unique() %>%
  dplyr::group_by(Type) %>%
  dplyr::slice_max(order_by = Zscore, n = 10, with_ties = F) %>%
  data.frame()
  
gg_umap <- ggplot(umap_df, aes(x=UMAP1, y=UMAP2, shape=Type, color=Type)) +
  geom_point(aes(size=Zscore), alpha=.5) +
  # geom_text(data = umap_labels, aes(x = UMAP1, y=UMAP2, label=Name))
  ggrepel::geom_label_repel(data = umap_labels,
                            aes(x=UMAP1, y=UMAP2, size=Zscore, color=Type, label=Name)) +
  theme_bw()

print(gg_umap)
```

## Construct similarity matrix

```{r Plot PD annotations, eval=FALSE} 
library(minet)# BiocManager::install("minet")  
#### Construct similarity matrix using mutual information #### 
# miM <- minet::build.mim(t(annot_mat))
miM <- minet::build.mim(t(umap_dist))
# Prune edges
# miM <- minet::aracne(as.matrix(umap_dist))
miM <- minet::clr(miM)
# sM <- BUS::gene.similarity(umap_res, measure="corr", net.trim = "none")
# pM <- BUS::gene.pvalue(umap_res, measure="MI", net.trim = "none")
# netM <- BUS::pred.network(pM$multi.perm.p.value, sM)
 
edges <- reshape2::melt(miM) %>% `colnames<-`(c("from","to","weight"))
edges <- edges[!is.na(edges$weight),]
edges <- edges[edges$weight>0,]
hist(edges$weight, 50) 
```

## Construct network

```{r, eval=FALSE} 
# %>% tibble::column_to_rownames("Type_ID")
net <- network::network(edges,  
                        # vertex.attr = data.frame(umap_df[!duplicated(umap_df$Type_ID),],  row.names = umap_df[!duplicated(umap_df$Type_ID),]$Type_ID)[unique(edges$from),],
                        matrix.type = "edgelist", 
                        layout = "fruchtermanreingold",
                        ignore.eval = F) 
# net.sum <- network::summary.network(net)
# ?gplot.layout {sna}	
netF <- ggnetwork::ggnetwork(net, 
                  # layout="circle",
                  layout = "fruchtermanreingold"
                  # cell.jitter = 0.75
                  )  %>%
  merge(umap_df, by.x = "vertex.names", by.y = "Type_ID")

# netF <- dplyr::group_by(netF, Type) %>% 
#   dplyr::slice_max(order_by = Zscore, n = 100, with_ties = F)
# dplyr::group_by(netF, Type) %>% count()
hubs <-  netF %>%  
  dplyr::mutate(Type_ID=paste(Type,ID,sep="_")) %>% 
  dplyr::group_by(Type, vertex.names, Name) %>% 
  dplyr::summarise(Type_ID, n=n(), x=max(x), y=max(y), Zscore=mean(Zscore)) %>%
  dplyr::group_by(Type) %>%
  dplyr::arrange(desc(n), desc(Zscore))  %>%
  dplyr::slice_max(order_by = n, n = 20, with_ties = F) %>%
    dplyr::relocate(Type_ID, 1) 
hubs%>% group_by(Type) %>% count() 
```

## Plot network 

### ggnetwork 

```{r, eval=FALSE} 
library(ggnetwork)
gg_net <- ggplot(netF, 
       aes(x = x, y = y, xend = xend, yend = yend, label=Name)) +
  geom_edges(aes(alpha=weight),# curvature = .3,
             # curvature = .3, ncp = 100
             color = "black"
             ) +
  geom_nodes(aes(size=Zscore, shape=Type, fill=Type, color=Type), alpha=.8) +
  # geom_nodetext(aes(color = Type, label = Name, size=Zscore),
                # fontface = "bold", nudge_y = .02, show.legend = F) +
  ggrepel::geom_label_repel(data = hubs, 
                           aes(x=x, y=y, label=Name, size=Zscore, color=Type), 
                           inherit.aes = F, alpha=.8, max.overlaps = 20) +
  theme_blank() 
print(gg_net)
```


### [ggraph](https://github.com/thomasp85/ggraph)

Has [edge bundling](https://www.r-graph-gallery.com/311-add-labels-to-hierarchical-edge-bundling.html)!....But can be quite slow with medium-large networks.

[Converting between hclust/networks.](https://www.gastonsanchez.com/visually-enforced/how-to/2014/06/29/Graph-from-dendrogram/)

```{r, eval=FALSE}
library(ggraph)
library(igraph)

hist(edges$weight)
edges <- subset(edges, from %in% hubs$vertex.names | to %in% hubs$vertex.names)
nodes <- annot_unique %>% dplyr::relocate(Type_ID, 1) %>% unique()
graph <- graph_from_data_frame(netF, 
                               directed = T,
                               # vertices = nodes[!duplicated(nodes$Type_ID),]
                               )
# graph <- graph_from_adjacency_matrix(umap_dist, weighted = T)
# phy <- umap_res$embedding %>% 
#   `colnames<-`(paste("UMAP",1:ncol(umap_res$embedding))) %>%
#   `row.names<-`(row.names(annot_mat)) %>%
#   dist() %>%
#   hclust()  %>%
#   ape::as.phylo()

# library(network)
# graph_net<- ape::as.network.phylo(phy,nodes)
# graph_net = graph.edgelist(phy$edge)
# graph_net <- ape::as.igraph.phylo(phy)
# igraph::vertex.attributes(graph_net) <- nodes
 
#  
# ggraph(net, layout = "dendrogram",circular=T) +
#    geom_node_point(alpha=.8) +
#    # geom_edge_link() 
#    geom_edge_diagonal() +
#     coord_fixed()
# 
# ggraph(net, layout = "treemap") + 
#   geom_node_tile(aes(fill = depth), size = 0.25)
#   
# ggraph(net, layout = 'circlepack') + 
#   geom_node_circle(aes(fill = depth), size = 0.25, n = 50) + 
#   coord_fixed()

#### Simple graph ####
ggg <- ggraph(graph, layout = "kk") + 
    geom_edge_link(aes(alpha=weight)) +
   geom_node_point(aes(color=Type, shape=Type, size=Zscore), alpha=.8)+
  geom_node_label(data=hubs, aes(label=Name, color=Type, size=Zscore),
                   alpha=.8, repel = T) + 
  theme_graph()
print(ggg)  
```

### [Edge bundling](https://www.r-graph-gallery.com/310-custom-hierarchical-edge-bundling.html) 

```{r Edge bundling, eval=FALSE}
# Create a graph of the flare class system
library(tidygraph)
# vertices <- umap_df %>% dplyr::select(name=Type_ID, size=Zscore, shortName=Type)
# edges <- reshape2::melt(miM) %>% `colnames<-`(c("from","to","weight"))
# edges <- edges[!is.na(edges$weight),]
# edges <- edges[edges$weight>0,]

tgraph <- tidygraph::tbl_graph(edges, nodes=nodes, node_key="Type_ID")

importFrom <- match(edges$from, nodes$Type_ID)
importTo <- match(edges$to, nodes$Type_ID)

# Use class inheritance for layout but plot class imports as bundles
gg_bundle <- ggraph(tgraph, 'kk') +
  geom_conn_bundle(aes(colour = stat(index)),
    data = get_con(importFrom, importTo),
    edge_alpha = 0.25, tension = 1
  ) +
  geom_node_point() +
  scale_edge_colour_distiller('', direction = 1, guide = 'edge_direction') +
  geom_node_point(aes(color=Type, shape=Type, size=Zscore), alpha=.8)+
  geom_node_label(data=hubs, aes(label=Name, color=Type, size=Zscore),
                   alpha=.8, repel = T) + 
  # coord_fixed() +
  ggforce::theme_no_axes()

```


# [Animated UMAP](https://rpubs.com/sbet/notebook_parliamentelections_municipalities) 
 
# Session info 
 
<details>

```{r}
utils::sessionInfo()
```

</details> 


